---
title: 3_Visualisation.ipynb
output: html_notebook
---

Basic installation/update:
```{r, eval=FALSE}
natmanager::install('natverse')
natmanager::install(pkgs = 'jefferis/vfbconnectr')
# update Python package
vfbconnectr::vc_install()
```


```{r, message=FALSE}
# load packages
library(vfbconnectr)
library(natverse)
library(dplyr)

# if Running on Rstudio server (==cloud) use plotly as visualisation default
#if(isTRUE(Sys.getenv('RSTUDIO_PROGRAM_MODE')=="server"))
options(nat.plotengine = 'plotly')
```

Set up connections to VFB APIs and VFB CATMAID server.

```{r}
vc = VfbConnect()
# Connect to the VFB CATMAID server hosting the FAFB data
catmaid::vfbcatmaid('fafb')
```


# Plotting
`nat` lets you plot neurons in 2d using, and in 3d using either `plotly` when
in a notebook environment or using a `rgl`-based 3D viewer when running on the
desktop. The `rgl` solution won't work in rstudio.cloud so we will focus on 2d
and `plotly` for 3d. See http://natverse.org/nat/articles/plotly.html for more details.


We'll use a FAFB neuron, retrieved using nat/rcatmaid as an example
This retrieves a CATMAID neuron object with skeleton id = 16 
(See the Mapping notebook for how to find ids)
```{r}
n = read.neuron.catmaid(16)
# just typing gives you information about the neuron
n
```

You can get more details like this:

```{r}
summary(n)
```

```{r}
# We can plot this in 2D
plot(n)
```

```{r}
plot(n, WithNodes = F)
```

```{r}
plot(n, WithNodes = F, col='gold')
```


```{r}
# in order to 
nclear3d()
plot3d(n, plotengine='plotly', soma=T)
```

**Navigation:**

* left click and drag to rotate (select "Orbital rotation" above the legend to make your life easier)
* mousewheel to zoom
* right-mouse + drag to translate
* click legend items (single or double) to hide/unhide. (FIXME)

**Customization:**

The above plots are very basic examples but there are a ton of ways to tweak things to your liking.
For a full list of parameters check out the docs for plot and plot3d.

Let's for example change the colors. In general, colors can be:

a string - e.g. "red" or
an rgb/rgba tuple as hex e.g. "#FF0000". 

```{r}
nclear3d()
plot3d(n, col='cyan')
```

```{r}
nclear3d()
plot3d(n, col='#FF0000')
```


#### Excercise

Float over the 

```{r}
# CATMAID neuron objects include connectors (synapses)  here pre-(red) and postsynapses (blue).
nclear3d()
plot3d(n, width=1000, WithConnectors = TRUE, col='cyan', plotengine='plotly') #
```

Now let's try an example with multiple neurons

```{r}
# Pulling (annotations on) neurons from a single dataset on CATMAID

batesids = catmaid_skids('Paper: Bates and Schlegel et al 2020') 
# Viewing first 10 neurons:
bates10=read.neurons.catmaid(batesids[1:10])
bates10
```

```{r}
nclear3d()
# Plotting multiple neurons, we get a default multi-color pallete
plot3d(bates10)
```

What if we want to use colour to group neurons in some way?

The following example illustrates colouring by lineage, where lineage is derived from standardised neuron names on FAFB CATMAID 


```{r}
# On CATMAID, we need to know something about the structure of names and parse them to get types:
da1=read.neurons.catmaid("name:Uniglomerular(.*?) DA1 ")
da1
```

```{r}
# look at the catmaid metadata for those neurons
da1[,]
```

```{r}
# make a new column in the attached metadata for lineages
da1[,'lineage']=stringr::str_match(da1[,'name'], '[adlv]+PN')
table(da1[,'lineage'])
plot(da1, col=lineage, WithNodes=F)
```

We can do the same thing using VFB, looking up the relevant terms on the web site and then using these to categorise CATMAID neurons

```{r}
# With VFB, we can query by type
vfb_type_2_skids <- function(vfb_type) {
    gen_short_form <- vfbconnectr:::vfb_connect$cross_server_tools$gen_short_form
    # Get IDs (short_forms) for instances of type
    urlids=vc$oc$get_instances(vfb_type, query_by_label=TRUE)
    da1_from_vfb = sapply(urlids, gen_short_form)
    
    # Find which neurons are in catmaid_fafb and return 
    da1_skid_lookup = vc$neo_query_wrapper$vfb_id_2_xrefs(da1_from_vfb, db='catmaid_fafb', reverse_return=TRUE)
    # Convert skids to ints 
    names(da1_skid_lookup)
}

da1_fafb = vfb_type_2_skids("'adult antennal lobe projection neuron DA1'")
da1_l_fafb = vfb_type_2_skids("'adult antennal lobe projection neuron DA1 lPN'")
da1_v_fafb = vfb_type_2_skids("'adult antennal lobe projection neuron DA1 vPN'")

#da1 = bates[is_da1]
```


```{r}
# add a new metadata column with the lineage calculate from FAFB data
da1[,'fafblineage']=""
da1[da1_l_fafb,'fafblineage']='lPN'
da1[da1_v_fafb,'fafblineage']='vPN'
da1[,c('skid', 'lineage', 'fafblineage')]
```


```{r}
plot(da1, col=fafblineage, WithNodes=F)
```

VFB allows us to pull neurons from multiple sources and display them in a single template space

```{r}
unlink('DA1', recursive = T)
DA1_manifest = vc$get_images_by_type("'adult antennal lobe projection neuron DA1'", template = 'JRC2018Unisex', image_folder = 'DA1', stomp=TRUE)
```

```{r}
DA1_manifest
```

We can then plot the neurons, colouring by data_source

```{r}
da1all=read.neurons("DA1/", df = DA1_manifest)
da1all[,'dataset']
```


```{r}
plot(da1all, col=dataset, WithNodes=F)
```

### Adding volumes

```{r}
# For a full list of volumes available on JRC2018Unisex, please see:
# https://v2.virtualflybrain.org/org.geppetto.frontend/geppetto?q=VFB_00101567,PaintedDomains

# Get an image (obj volume) of the mushroom body calyx on JRC2018Unisex: 
unlink('MB', recursive = TRUE)
MB = vc$get_images_by_type("'mushroom body calyx'", template = 'JRC2018Unisex', image_folder='MB', image_type='obj', stomp=TRUE) # TODO - check for man and pull if there.
MB
```

```{r}
mbca = read.neurons('MB/CA_on_JRC2018Unisex_adult_brain.obj')
boundingbox(mbca)
```


```{r}
nclear3d()
plot3d(da1all[1:10],col=dataset)
# FIXME can't seem to shade in plotly
# @Sri ideas?
wire3d(mbca[[1]], col='grey')
```
