---
title: "xform example"
output: html_notebook
---

This notebook is going to give an example of transforming neurons between different brain spaces. In particular we will use FAFB and hemibrain data.

```{r, message=FALSE}
library(natverse)
# for the cloud we need to use plotly for 3D plots 
# see http://natverse.org/nat/articles/plotly.html for details
options(nat.plotengine = 'plotly')
```

Quick check to make sure everything is working ok. 
```{r}
plot(Cell07PNs[1:2])
```

OK so the basic check looks good.

Now let's fetch some neurons from [vfb catmaid](http://catmaid.virtualflybrain.org/):

```{r}
vfbconn=vfbcatmaid("fafb")
dl1.fafb=read.neurons.catmaid('name:DL1.*PN', conn = vfbconn)
```

We can make a 2d plot like so:
```{r}
# With Nodes suppresses plotting all the nodes. 
plot(dl1.fafb[[1]], WithNodes = F)
```

Although this is not a problem for the rgl package used for visualisation on the 
desktop, it may be a good idea to simplify the neurons for display in the browser

```{r}
simplifyn <- function(x, stepsize=1, twigsize=2) {
  # crude way to guess if units are in nm or microns
  maxdim=max(diff(boundingbox(dl1.fafb)))
  units=if(maxdim>2000) "nm" else "um"
  stepsize=ifelse(units=="nm", stepsize*1000, stepsize)
  twigsize=ifelse(units=="nm", twigsize*1000, twigsize)
  x1=prune_twigs(x, twig_length=twigsize)
  x2=resample(x, stepsize)
  x2
}
```


```{r}
nvertices(dl1.fafb)
nvertices(simplifyn(dl1.fafb))
summary(dl1.fafb)
summary(simplifyn(dl1.fafb))
```
OK, so let's get to work
```{r}
plot3d(simplifyn(dl1.fafb))
```


OK so now we've got some FAFB neurons, but we'd also like to get some hemibrain ones.

```{r}
library(neuprintr)
neuprint_login()
```

```{r}
neuprint_get_meta('DL1_adPN')
```

The LHS neurons will be really be truncated, so let's stick with RHS.
```{r}
# note how we specify neurons with a certain name
dl1.fib=neuprint_read_neurons("name:DL1_adPN_R")
```

```{r}
dl1.fib
```

Small hack. Switch to rgl with webgl to plot synapses

```{r}
rgl::setupKnitr()
```

```{r, webgl=T}
options(nat.plotengine = 'rgl')
nclear3d()
# FIXME @Greg the simplifyn removes the connectors, so why do we plot connectors here?
plot3d(simplifyn(dl1.fib), WithConnectors=T, col='grey')
options(nat.plotengine = 'plotly')
```

```{r}
library(nat.jrcbrains)
#if you are in the r-cloud use an already downloaded folder like below and register it..
options(nat.jrcbrains.regfolder='/cloud/project/bridging/nat.jrcbrains/')
register_saalfeldlab_registrations()
```

```{r}
# this correction factor converts raw 8nm voxels 
dl1.fib.fafb=xform_brain(dl1.fib*(8/1000), sample='JRCFIB2018F', ref='FAFB14')
```


```{r}
plot(dl1.fafb, WithNodes=F)
plot(dl1.fib.fafb, col='grey', add=T, WithNodes=F)

```

```{r}
nclear3d()
plot3d(simplifyn(dl1.fafb))
plot3d(simplifyn(dl1.fib.fafb), col='grey')
```

```{r}
dl1.fafb[,'brain']='FAFB'
dl1.fib.fafb[,'brain']='hemibrain'
dl1.all=c(dl1.fafb, dl1.fib.fafb)
```

```{r}
nclear3d()
plot3d(simplifyn(dl1.all), col=brain, soma=T)
plot3d(FAFB)
```

We can also move neurons from the left to right of FAFB. It's convenient to 
add a column to keep track of which neurons are on the left and right.

```{r}
# find mean X axis position
meanx=sapply(dl1.all, function(n) mean(xyzmatrix(n)[,1]))
meanx
dl1.all[,'side']=ifelse(meanx>500e3, "L", "R")
```

We can use subset to mirror only the LHS neurons
```{r}
dl1.all.r=mirror_fafb(dl1.all, subset = dl1.all[,'side']=="L")
```

```{r}
nclear3d()
plot3d(simplifyn(dl1.all.r), col=side)
plot3d(FAFB)
```

```{r}
nclear3d()
dl1.all.r.simp = simplifyn(dl1.all.r)
plot3d(dl1.all.r.simp, col='grey', WithNodes =T)
```
